---
title: 'Introducing the drake package'
author: "Hans Van Calster"
date: "13 februari 2020"
subtitle: ""
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: ["default", "extra.css", "default-fonts"]
    chakra: libs/remark-latest.min.js
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
options(servr.daemon = TRUE)
library(knitr)
opts_chunk$set(echo = FALSE)

library(tidyverse)
library(drake)
```



```{r eval=FALSE, include=FALSE}
xaringan::summon_remark(version = "latest", to = "libs/")
```

## What's this all about?

An evolution from:

- r-scripts in `.R` files

--

- to `Rmarkdown` and `bookdown`
    - scripts to documented scripts (literate programming)
    - learn to avoid copy-paste
        - write functions
        - use `purrr`
        - use parent-child documents

--

But data-analysis and reporting is better decoupled

--

- markdowns with heavy calculations
    - long runtime
        - chunk option `cache = TRUE`, but can be tricky
    - (too) big objects in the R environment

---

## Enter `drake`

- analysis pipeline comes first and is followed by reporting
    - reporting can be integrated in the 'pipeline' as a last step (`knitr_in()` function)
    - or completely separated (for instance in a different R project)

- stimulates to work with functions as much as possible


--

Short intro to the `drake` package: https://books.ropensci.org/drake/index.html#short-version

```{r out.height="20%", out.width="20%", fig.align='center'}
include_graphics(path = "drake_hex_logo.svg")
```


---

## Simple workflow

1. `drake_plan()`

2. `make()`

3. in case of errors `diagnose()`

4. access the targets with `readd()` or `loadd()` and check them

5. change the plan in 1.

6. go to 2.

---

## Project structure for more complicated workflows

````
|- prj_drake
|    |- prj_drake.Rproj
|    |- run.R
|    |- _drake.R
|    |- R
|    |    |- plan.R # the workflow containing all targets that need to be calculated
|    |    |- functions_*.R # all functions that are needed
|    |    |- packages.R # all packages that are needed
````

Example _drake.R:

```{r echo=TRUE, eval=FALSE}
source("R/packages.R")
source("R/functions.R")
source("R/plan.R")
# options(clustermq.scheduler = "multicore") # optional parallel computing
drake_config(plan, verbose = 2)
```

---

## Continued

Example run.R:

```{r echo=TRUE, eval=FALSE}
library(drake)
r_outdated()

r_make()

r_outdated()

library(visNetwork) # so the graph renders
r_vis_drake_graph()
```

`r_make()` and other `r_*` will call _drake.R

---
class: middle, center



<iframe src="florabank_vis_graph.html" width="100%" height="80%" id="igraph"  scrolling="no" seamless="seamless" frameBorder="0" > </iframe>

## Scale-up your work

---

## Static branching

```{r echo=TRUE, eval=FALSE}
drake_plan(
  trace = TRUE,
  # all combinations 1a 1b 2a 2b
  crossed_targets = target(
    my_function(..., factor_A, factor_B),
    transform = cross(
      factor_A = c(1, 2),
      factor_B = c("a", "b")
    ),
    # rowwise combinations 1a and 2b
    mapped_targets = target(
    my_function(..., factor_A, factor_B),
    transform = map(
      factor_A = c(1, 2),
      factor_B = c("a", "b")
    )
  )
)

```

---


## Dynamic branching

>Dynamic branching is an interface to declare new targets while `make()` is running. It lets you create more compact plans and graphs, it is easier to use than static branching, and it improves the startup speed of `make()` and friends.

https://books.ropensci.org/drake/dynamic.html#which-kind-of-branching-should-i-use



---

## What else?

- output formats

- `clean()` function

- multiple caches

---

## What else?

- integration with git
    - the `.cache` folder is `gitignored`
    - `make(cache_log_file = TRUE)` -> creates a csv log file with hashes for each target. This file should be put under version control
    - drake also has a recovery feature
    
- integration with `docker`: 
    - drake outside docker: https://github.com/wlandau/drake-examples/tree/master/Docker-psock
    - drake inside docker: https://gitlab.com/ecohealthalliance/drake-gitlab-docker-example

- integration with `renv`:https://github.com/rstudio/renv
    - ensures everyone will work with the same versions of R packages

- integration in a package: https://books.ropensci.org/drake/projects.html#workflows-as-r-packages

---

## Resources

- check out https://github.com/ropensci/drake

















